Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-06-21 draft.tex

@@

\subsection{Supernova hosts \label{subsec:incompletenss_SN_hosts}}
A further limiting factor in the SN host sample is the requirement of a spectroscopic redshift. The majority of SN host spectroscopic redshifts in DES are provided by the dedicated follow-up survey OzDES, for which the limiting magnitude is around 24 to 24.5 mag in the $r$ band \citep{Lidman2020}. The rate at which a host redshift is successfully measured given an apparent magnitude has been extensively modelled by \citet{Vincenzi2020} who provide the spectroscopic redshift efficiency $\epsilon_{z_{\mathrm{spec}}} (m^{\mathrm{host}}_{r,i})$ as a function of host $r$-band magnitude, host galaxy colour, and the year in which the SN was discovered in order to allow for a longer possible spectroscopic exposure time for hosts of SNe discovered earlier in the survey. The effects of the host galaxy spectroscopic efficiency are displayed in the upper panel of Fig. \ref{fig:mass_hist_hosts_field}, where the normalised histogram weighted by the efficiency is skewed towards lower masses when compared to the unweighted distribution. So that we reduce any bias towards SNe in bright hosts that are easier to obtain spectroscopic redshifts for, the host galaxy spectroscopic redshift efficiency is multiplied by that derived from the SN detection efficiency (Section \ref{subsec:incompleteness_SNe}) to arrive at the final efficiency for the $i$th SN:
\begin{equation}
    \eta_{\mathrm{SN},i} = \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i},x_{1,i},c_i)\times \epsilon_{z_{\mathrm{spec}}} (m^{\mathrm{host}}_{r,i})\,.
\end{equation}

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_x1_fitjoint.png}
    \caption{As per Fig. \ref{fig:rate_fitted_split_x1} but with the data fit simultaneously by a single two-component DTD, with fixed slope $\beta = -1$ and a prompt fraction that is 1 at short delay times, 0 at large delay times, and a linear function of time between $t_1 =0.65$~Gyr (where tardy SNe switch on) and $t_2=0.85$~Gyr (where prompt SNe turn off).%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:rate_fitted_split_x1_2c_DTD}}
\end{figure}
The SN Ia rate as a function of stellar mass for SNe split by stretch is shown in Fig. \ref{fig:rate_fitted_split_x1}. The evolution of the rate with stellar mass is significantly different for the two sub-samples: high-stretch SNe dominate the rate at low stellar mass but tail off in the higher mass galaxies, while low-stretch SNe are subdominant in low mass galaxies but display a much steeper dependence on stellar mass and make up the vast majority of SNe in high-mass hosts. These measurements reflect the observed correlations of $x_1$ with stellar mass and sSFR (and by inference, stellar age) as been seen by \citet{Rigault2013,Graur2017,Rigault2018,Rose2019,Nicolas2020,Rose2021}. The corresponding DTD fit results in a steep power-law $\beta =-1.70^{+0.19}_{-0.10}$ for high-stretch SNe indicative of a population of predominantly prompt SNe, and a much shallower decay slope ($\beta = -0.79 \pm 0.08$) for low-stretch SNe representing a much more delayed population. The difference in $\beta$ is significant at a level of $3.8\sigma$. This result provides an intriguing confirmation of that observed in 101 SDSS SNe Ia by \citet{Brandt2010} who found that low-$x_1$ SNe displayed much longer delay times than those with high $x_1$.

>>>>

\begin{figure}
    \centering
    %\includegraphics[width=0.5\textwidth]{figs/SN_efficiencies.png}
    \includegraphics[width=0.5\textwidth]{figs/eff_z.png}
    \caption{SN host galaxy spectroscopic efficiencies as a function of redshift $z$. Efficiency curves are grouped by shallow (\{F\}12) and deep (\{F\}3) fields, and averaged over the 5 years of the survey. The lower and upper redshift limits for this analysis are indicated by $z_{\mathrm{min}}$ and $z_{\mathrm{max}}$ respectively.
    \label{fig:eff_z}}
\end{figure}

@@

\subsection{Supernova hosts \label{subsec:incompletenss_SN_hosts}}
A further limiting factor in the SN host sample is the requirement of a spectroscopic redshift. The majority of SN host spectroscopic redshifts in DES are provided by the dedicated follow-up survey OzDES, for which the limiting magnitude is around 24 to 24.5 mag in the $r$ band \citep{Lidman2020}. The rate at which a host redshift is successfully measured given an apparent magnitude has been extensively modelled by \citet{Vincenzi2020} who provide the spectroscopic redshift efficiency $\epsilon_{z_{\mathrm{spec}}} (m^{\mathrm{host}}_{r,i})$ as a function of host $r$-band magnitude, host galaxy colour, and the year in which the SN was discovered in order to allow for a longer possible spectroscopic exposure time for hosts of SNe discovered earlier in the survey. To assess how $\epsilon_{z_{\mathrm{spec}}} (m^{\mathrm{host}}_{r,i})$ translates into an efficiency as a function of redshift $\epsilon_{z_{\mathrm{spec}}} (z)$ within the redshift limits of our survey, we use a simulation of host galaxies as described in \citet{Vincenzi2020}. The simulation uses galaxy photometry from DES Science Verification images, and then selects them as potential SN hosts according a weighting driven by previously measured SN rate - host galaxy relationships. We note that this is for display purposes only, so the relationships built into the simulation are not propagated to the final rate measurements. For bins in redshift, we find the mean efficiency for measuring a spec-$z$ of all galaxies in the bin via $\epsilon_{z_{\mathrm{spec}}} (m^{\mathrm{host}}_{r,i})$. The resulting $\epsilon_{z_{\mathrm{spec}}} (z)$ curves are displayed in Fig. \ref{fig:eff_z} and show a high spectroscopic efficiency within the redshift limits of this analysis.

The effects of the host galaxy spectroscopic efficiency are displayed in the upper panel of Fig. \ref{fig:mass_hist_hosts_field}, where the normalised histogram weighted by the efficiency is skewed towards lower masses when compared to the unweighted distribution. So that we reduce any bias towards SNe in bright hosts that are easier to obtain spectroscopic redshifts for, the host galaxy spectroscopic redshift efficiency is multiplied by that derived from the SN detection efficiency (Section \ref{subsec:incompleteness_SNe}) to arrive at the final efficiency for the $i$th SN:
\begin{equation}
    \eta_{\mathrm{SN},i} = \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i},x_{1,i},c_i)\times \epsilon_{z_{\mathrm{spec}}} (m^{\mathrm{host}}_{r,i})\,.
\end{equation}

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_x1_fitjoint.png}
    \caption{\textit{Upper}: Posterior samples of the fraction of "prompt" SNe resulting from the two-population fit, displayed as a function of host galaxy stellar mass. 
    \textit{Lower}: As per Fig. \ref{fig:rate_fitted_split_x1} but with the data fit simultaneously by a single two-component DTD, with fixed slope $\beta = -1$ and a prompt fraction that is 1 at short delay times, 0 at large delay times, and a linear function of time between $t_1 =0.65$~Gyr (where tardy SNe switch on) and $t_2=0.85$~Gyr (where prompt SNe turn off).%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:rate_fitted_split_x1_2c_DTD}}
\end{figure}
The SN Ia rate as a function of stellar mass for SNe split by stretch is shown in Fig. \ref{fig:rate_fitted_split_x1}. The evolution of the rate with stellar mass is significantly different for the two sub-samples: high-stretch SNe dominate the rate at low stellar mass but tail off in the higher mass galaxies, while low-stretch SNe are subdominant in low mass galaxies but display a much steeper dependence on stellar mass and make up the vast majority of SNe in high-mass hosts. These measurements reflect the observed correlations of $x_1$ with stellar mass and sSFR (and by inference, stellar age) as been seen by \citet{Rigault2013,Graur2017,Rigault2018,Rose2019,Nicolas2020,Rose2021}. The corresponding DTD fit results in a steep power-law $\beta =-1.70^{+0.19}_{-0.10}$ for high-stretch SNe indicative of a population of predominantly prompt SNe, and a much shallower decay slope ($\beta = -0.79 \pm 0.08$) for low-stretch SNe representing a much more delayed population. The difference in $\beta$ is significant at a level of $3.8\sigma$. This result provides an intriguing confirmation of that observed in 101 SDSS SNe Ia by \citet{Brandt2010} who found that low-$x_1$ SNe displayed much longer delay times than those with high $x_1$.
